---
title: "Dashboard - Marketing Campaign Response"
author: Anna Bukharova
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}
# install.packages("flexdashboard")
library(flexdashboard)
library(lubridate)
library (ggplot2)
library(magrittr) 
library(dplyr)
library (coin)
library(rpart.plot)
library(readr)
library(pROC)

data = read_csv("~/Documents/r files/ml_project1_data.csv")
data$week = wday(data$Dt_Customer, label = TRUE)
data$Response = as.factor(data$Response)
```

```{r}
# Income groups
data3 = data %>% filter(!is.na(Income))
mean_income = mean(data3$Income)
data3 = mutate(data3,
  income_group = factor(case_when(
    Income < mean_income ~ "Below average",
    TRUE ~ "Above average"
  ))
)

# Model
set.seed(888)
train_index = sample(1:nrow(data), 
                      size = 0.8 * nrow(data))
train = data[train_index, ]
test  = data[-train_index, ]

tree = rpart(
  Response ~ Income + NumWebPurchases + Recency,
  data = train, method = "class",
  weights = ifelse(train$Response == 1, 5, 1)
)
pred_class = predict(tree, newdata = test, 
                      type = "class")
pred_prob  = predict(tree, newdata = test, 
                      type = "prob")
accuracy   = round(mean(pred_class == test$Response), 2)
roc_obj    = roc(test$Response, pred_prob[, 2], 
                  quiet = TRUE)
auc_val    = round(auc(roc_obj), 2)
```

Page 1: Customer research {data-icon="fa-users"}
===

Column {data-width=500}
---

### Income level affects campaign response

```{r}
plot_income = data3 %>%
  filter(Response == 1) %>%
  group_by(income_group) %>%
  summarise(count = n()) %>%
  mutate(pct = round(count / sum(count) * 100, 1))

ggplot(plot_income, 
       aes(x = income_group, y = count)) +
  geom_bar(stat = "identity", fill = "#104E8B") +
  geom_text(aes(label = paste0(count, 
                               " (", pct, "%)")),
            vjust = -0.5, size = 3.5) +
  labs(x = "Income group", 
       y = "Number of responses") +
  theme_classic()
```

### Income × Response

```{r}
valueBox("p < 0.001 — significant", 
         icon = "fa-check-circle",
         color = "success")
```

Column {data-width=500}
---

### Recently active customers respond more

```{r}
ggplot(data, aes(x = Response, y = Recency)) +
  geom_boxplot(fill = "#BFEFFF") +
  labs(x = "Campaign response (0 = No, 1 = Yes)",
       y = "Days since last purchase") +
  theme_classic()
```

### Recency × Response

```{r}
valueBox("p < 0.001 — significant",
         icon = "fa-check-circle",
         color = "success")
```

Page 2: Purchase Behaviour {data-icon="fa-shopping-cart"}
===

Column {data-width=500}
---

### Web purchases associated with higher response

```{r}
ggplot(data, aes(x = Response, 
                 y = NumWebPurchases)) +
  geom_boxplot(fill = "#BFEFFF") +
  labs(x = "Campaign response (0 = No, 1 = Yes)",
       y = "Number of web purchases") +
  theme_classic()
```

### Web purchases × Response

```{r}
valueBox("p < 0.001 — significant",
         icon = "fa-check-circle",
         color = "success")
```

Column {data-width=500}
---

### In-store purchases show no significant pattern

```{r}
ggplot(data, aes(x = Response, 
                 y = NumStorePurchases)) +
  geom_boxplot(fill = "#BFEFFF") +
  labs(x = "Campaign response (0 = No, 1 = Yes)",
       y = "Number of store purchases") +
  theme_classic()
```

### Store purchases × Response

```{r}
valueBox("p = 0.05 — not significant",
         icon = "fa-times-circle",
         color = "warning")
```

Page 3: Predictive Model {data-icon="fa-chart-line"}
===

Column {data-width=600}
---

### Decision tree: key predictors of response

```{r}
rpart.plot(tree, 
           main = "Response prediction tree")
```

Column {data-width=400}
---

### Model accuracy

```{r}
valueBox(paste0(accuracy * 100, "%"),
         icon = "fa-bullseye",
         color = "primary")
```

### Area under ROC curve (AUC)

```{r}
valueBox(auc_val,
         icon = "fa-chart-area",
         color = ifelse(auc_val >= 0.7, 
                        "success", "warning"))
```


### Top targeting recommendation

```{r}
valueBox("Target: high income + recent + digital",
         icon = "fa-crosshairs",
         color = "success")
```

